---
title: "Progress Report - Lab 4"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(rvest)
install.packages("Lahman")
library(Lahman)
```

# Data Background

The Lahman package contains historical Hall of Fame data, including some data for 2026.

I will compare my scraped data with the Lahman dataset to see what matches and what is different.

# Data Source

I am using Baseball Reference to scrape Hall of Fame 2026 data.

Link: https://www.baseball-reference.com/awards/hof_2026.shtml

```{r}
# Scrape the data

url <- "https://www.baseball-reference.com/awards/hof_2026.shtml"
page <- read_html(url)
tables <- page %>% html_table(fill = TRUE)
tables
```

```{r}
tables[[1]]
tables[[2]]
```

```{r}
# Step 3: Select correct table
hof_data <- tables[[2]]
head(hof_data)
```

# Observations
- The webpage returned multiple tables.
- The first table is not relevant for Hall of Fame voting.
- The second table (tables[[2]]) contains the Hall of Fame voting data.
- This table includes player names, votes, and percentage of votes.

```{r}
# Step 4: Compare with Lahman data

lahman_data <- HallOfFame %>%
  filter(yearID == 2026)

lahman_data
```

# Comparison
- The Lahman dataset has fewer rows compared to the scraped data.
- Lahman includes official voting records, while the scraped data includes the full ballot.
- The variables are different between the two datasets.
- This explains why the datasets do not match exactly.


