---
title: "Progress Report"
author: "Pablo Leguizamo"
date: "2026-5-4"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(rvest)
library(readr)
library(Lahman)
```

# Data Source

I am using Baseball Reference to scrape the 2026 Hall of Fame voting data.

Link: https://www.baseball-reference.com/awards/hof_2026.shtml

```{r}
url <- "https://www.baseball-reference.com/awards/hof_2026.shtml"
html <- read_html(url)
tables <- html_table(html)
length(tables)
```

```{r}
data <- tables[[1]]
head(data)
```

- The webpage had more than one table.
- The first table has the BBWAA Hall of Fame voting results.
- The names of the columns were in the first row, so I fixed the column names.

```{r}
actual_col_names <- data[1, ]
colnames(data) <- actual_col_names
data <- data[-1, ]

head(data)
```

# Clean Data

I cleaned the names, changed votes to numbers, and kept the columns I needed.

```{r}
head(HallOfFame, 3)
```

```{r}
data$Name <- gsub("HOF", "", data$Name)
data$Name <- gsub("X-", "", data$Name)
data$Name <- trimws(data$Name)
data$Name <- iconv(data$Name, from = "UTF-8", to = "ASCII//TRANSLIT")
data$Name <- trimws(data$Name)
data$Votes <- as.numeric(data$Votes)
data$percent_vote <- readr::parse_number(data$`%vote`)

data$yearID <- 2026
data$inducted <- ifelse(data$percent_vote >= 75, "Y", "N")

hof_2026 <- data.frame(
  Name = data$Name,
  yearID = data$yearID,
  Votes = data$Votes,
  percent_vote = data$percent_vote,
  inducted = data$inducted
)

head(hof_2026, 3)
```

# Comparison

```{r}
lahman_2026 <- HallOfFame[HallOfFame$yearID == 2026, ]
lahman_2026
```

- The scraped data has the full BBWAA ballot from Baseball Reference.
- Lahman is organized differently and uses player IDs.
- My data keeps the name, year, votes, percent vote, and inducted result.
